Noninvasive dynamic output-feedback stabilization of multistable systems

### 2024-12-12 Paper draft/backup_1.tex
Rename forlders

		\author[First]{Samuele Zoboli} 		
        \author[Second]{Andreu Cecilia}
	\author[First]{Sophie Tarbouriech}

		\address[Second]{Universitat Polit\`ecnica de Catalunya, Avinguda Diagonal, 647, 08028 Barcelona, Spain, (e-mail: andreu.cecilia@upc.edu)}
  \address[First]{LAAS-CNRS, Universit\'e de Toulouse, CNRS, Toulouse, (e-mail: samuele.zoboli@laas.fr, sophie.tarbouriech@laas.fr)}

@@

\section{Problem formulation}
We consider nonlinear systems of the form
\begin{equation}\label{eqn:original_system}
\begin{aligned}
\dot x &=   A x  + D\phi(Ex) + Bu,  \\
 y&= Cx+ G\phi(Ex),
\end{aligned}
\end{equation}
where $x\in\RR^{n_x}$ is the state vector,  $u\in\RR^{n_u}$ the controlled input, $y\in\RR^{n_y}$ the measured outputs and $\phi:\RR^{n_v}\to\RR^{n_v}$ is a sufficiently smooth vector field. Additionally, to recover efficiently solvable conditions, we assume the existence of known matrices $R=R^\top, Q=Q^\top$ and $S$ such that 
\begin{equation}\label{eqn:generic_abstraction}
    \pmx{\Id_{n_x}\\\pderiv{\phi}{x}(v)}^\top\pmx{R&S^\top\\S&Q}\pmx{\Id_{n_x}\\\pderiv{\phi}{x}(v)}\preceq 0 
\end{equation}
holds for all $v\in\cV\subseteq \RR^{n_v}$ where $\cV\defeq\{v\in\RR^{n_v}:v=E x, x\in\cX\subseteq \RR^{n_x}\}$. We recall that such a class of nonlinearities includes common and interesting family of functions, such as (shifted) monotonic and differentially sector-bounded ones \citep{zoboli2024quadratic}. 
Finally, we assume that in the unforced case, namely when $u=0$, system \eqref{eqn:original_system} presents a set of fixed points $\cF_x^\star$. More formally, we assume the existence of a nonempty set of equilibria
\begin{equation}\label{eqn:eq_points}
 \cF_x^\star := \{x\in\RR^{n_x} \ | \ Ax+D\phi(E x)= 0\}.
\end{equation}
We highlight that $\cF^\star$ can include multiple elements, that is, system \eqref{eqn:original_system}  may present multiple equilibrium points for the unforced case $u=0$.

We aim to design a feedback control policy that makes (at least) one point $x^\star\in\cF_x^\star$ asymptotically stable. In other words, we want to make the set $\cF_x^\star$ attractive and forward-invariant.

One possibility is to reformulate the problem as a standard stabilization one, see e.g. \citep{isidori2017stabilization}. Differently put, the problem could be solved by selecting a specific point  $x^\star\in\cF_x^\star$ and designing a control law making $x^\star$ globally asymptotically stable. However, we highlight that the considered problem relaxes the more common stabilization problem in three directions. First, the goal is not to stabilize the system at a specific point in $\cF_x^\star$, we only need the system to converge to $\cF_x^\star$. Second, the proposed objective allows for the existence of multiple attractive equilibrium points for the closed-loop system.  Third, the set of equilibrium points $\cF_x^\star$ may be unknown to the designer. This third point is particularly relevant for feedback policies that include an integral action, e.g. \citep{Astolfi2017,GIACCAGLI2023111106}, which often require an unknown non-zero steady-state control signal to be generated by the integral component.
Another possibility is to invoke set-stability approaches, e.g. \citep{SHIRIAEV20001709}. However, we remark that $\cF_x^\star$ may be disconnected. Moreover, its elements are given by the (possibly unknown) equilibria of the unforced system dynamics. These facts impose additional restrictions that can impact the effectiveness and applicability of set-stability approaches.

Given the abovementioned constraints, in what follows, we present a solution that does not require knowledge of the open-loop equilibria and guarantees their stability via $2$-contraction \citep{Michael1995}. Our approach is based on the recent findings in \citep{ cecilia2023generalized, zoboli2024dynamic} and inspired by the techniques of derivative control and washout filters.

\section{Main result}
\subsection{Proposed approach}
We start by highlighting a necessary property controllers must satisfy to guarantee the open-loop equilibria are preserved by the closed loop.
From \eqref{eqn:original_system}, it is clear that for the closed-loop system to stabilize at  $x^\star\in\cF_x^\star$ the control input must vanish at such point.
% dynamic output-feedback law $u=\alpha(y,\xi)$ with state $\xi\in\RR^{n_\xi}$ must satisfy 
% \begin{equation}\label{eqn:noninvasive_cond}
% \alpha(Cx^\star+G\phi(Ex^\star),\xi^\star) = 0, \qquad \forall x^\star \in \cF^\star,
% \end{equation}
% where $\xi^\star$ is the equilibrium state of the controller corresponding to $x^\star$.
We will denote a control policy that satisfies such a property as \textit{noninvasive}, e.g. \citep{ANGELI2024152}. 
\begin{definition}[Noninvasive controller]
Consider the nonlinear system 
\[
\dot x = f(x,u), \qquad y =h(x),
\]
with $x\in\RR^{n_x}, y\in\RR^{n_y}$ and $u\in\RR^{n_u}$. Let 
\[
\cF^\star\defeq\{x\in\RR^{n_x}:f(x,0)=0\}.
\]
The dynamic output-feedback law $u=\alpha(y,\xi)$  with state $\xi\in\RR^{n_\xi}$ is said to be \textit{ noninvasive} if 
\begin{equation}\label{eqn:noninvasive_cond}
\alpha(h(x^\star),\xi^\star) = 0, \qquad \forall x^\star \in \cF^\star,
\end{equation}
where $\xi^\star$ is the controller equilibrium associated to $x^\star$.
\end{definition}

In this paper, we propose ensuring such a requirement via integral action. More specifically, we extend the dynamics of \eqref{eqn:original_system} as follows
\begin{equation}\label{eqn:extended_dyn}
\begin{aligned}
\pmx{\dot x\\ \dot z}&= \pmx{A & 0\\0&0} \pmx{x\\ z} + \pmx{B\\\Id_{n_u}} u + \pmx{D\\0} \phi\left(\pmx{E&0} \pmx{x\\ z} \right),  \\
 \pmx{y\\z}&= \pmx{C & 0\\0 & \Id_{n_u}}\pmx{x\\ z} + \pmx{G\\0} \phi\left(\pmx{E&0}\pmx{x\\ z} \right).
\end{aligned}
\end{equation}
At any equilibrium point $(x^\star,z^\star)$ the $z$-dynamics impose $u=0$, and thus $x^\star \in \cF_x^\star$.

@@

holds for all $\eta\in\cS\subseteq \RR^{n_\eta}$ with $\cS\defeq\{\eta\RR^{n_\eta}: \bE\eta\in\cV\}$, see \citep[Section 2.5]{zoboli2024dynamic}. 

@@

    D_e \defeq  \pmx{D\\0}, \qquad G_e \defeq  \pmx{G\\0}, \quad 
        \Phi \defeq \pmx{\Pi_2 & \Id_{n_\xi}\\ \Id_{n_\xi} & \Pi_1},

@@

        \widehat A &\defeq\pmx{A_e\Pi_2+B_e \widehat L && A_e + B_e  \widehat K C_e \\ \widehat M && \Pi_1 A_e+ \widehat N  C_e }, \quad  
        \widehat D \defeq \pmx{D_e +B_e \widehat K G_e\\ \Pi_1 D_e  + \widehat N G_e}, 
        \quad  \widehat R \defeq \pmx{ \widehat R_{11} &  \widehat R_{12}^\top\\
        \star & \widehat R_{22}}, \quad \widehat S^\top \defeq \pmx{\widehat S_1^\top \\ \widehat S_2^\top},
        &
        \pmx{\widehat R_{11} - \overbar R& \widehat R_{12}^\top - \alpha \Pi_2 R^\top & \widehat S_{1}^\top-\alpha \Pi_2 S^\top\\
         \star & \widehat R_{22}-\alpha R & \widehat S_{2}^\top -\alpha S^\top \\
         \star&\star& \widehat Q-\alpha Q} +
         \He{\pmx{\Upsilon_1\\\Upsilon_2\\\Upsilon_3}\pmx{\pderiv{\phi}{x}(v)& 0 & -\Id_{n_x}}}\preceq 0 , \qquad \forall v \in \cV,

@@

   \He{A_e \Pi_2+B_e \widehat L} -2\beta_1 \Pi_2 - \widehat R_{11} && A_e+B_e \widehat K C_e  + \widehat M^\top-2\beta_1 \Id_{n_\xi} -  \widehat R_{12}^\top && D_e +B_e \widehat K G_e -  \widehat S_1^\top\\
   \star && \He{ \Pi_1 A_e +  \widehat N  C_e } -2\beta_1 \Pi_1   -\widehat R_{22}&& \Pi_1 D_e  + \widehat N G_e -\widehat S_2\\
   \star &&\star&& - \widehat Q

@@

    \star && \star& &\He{\Gamma_3 \widehat D} - \widehat Q

@@

and let \eqref{eqn:generic_abstraction_ctrl} hold for all  $x\in\cF$  with given matrices $R, S, Q$. Select $Z\preceq 0$ such that $Z\preceq R$ and assume there exist $ \overbar R_{11}=\overbar R^\top,$ $ \widehat R_{11}=\widehat R_{11}^\top,$ $ \widehat R_{12},$ $ \widehat R_{22}=\widehat R_{22}^\top,$ $ \widehat S_1,$ $ \widehat S_2,$ $ \widehat Q=\widehat Q^\top,$ scalars $\alpha>0,\beta_0\in\RR$, $\beta_1<0$,  symmetric matrices $\Pi_1=\Pi_1^\top, \Pi_2=\Pi_2^\top \in \RR^{n_\xi\times n_\xi}$, matrices $\widehat K \in \RR^{n_u\times n_y}$, $\widehat L\in \RR^{n_u\times n_\xi}$, $\widehat M\in \RR^{n_\xi\times n_\xi}$, $\widehat N \in \RR^{n_\xi\times n_\xi}$,  matrices $\Gamma_1, \Gamma_2, \Gamma_3$  and $\Sigma=\Sigma^\top\succ 0$  such that 

@@

        \item The matrix $ R_Z \defeq \smallpmat{\overbar R && \alpha \Pi_2Z\\ \star && \alpha Z}$ is negative semidefinite;
         \item Inequality \eqref{eqn:2_contraction_BMI}\footnotemark[1] holds with $\beta_0+\beta_1<0$;

@@

     there exist square invertible matrices $U,V\in \RR^{n_\xi\times n_\xi}$ such that $UV^\top = \Id_{n_\xi}- \Pi_1\Pi_2$

@@

         \pmx{N& F}&= U^{-1} (\widehat N - \Pi_1 B \widehat K) ,\\
           L & = (\widehat L - \widehat K C \Pi_2) V^{-\top},\\
         M&=U^{-1}\!\pmx{\Id_{n_\xi}\\-\Pi_1}^{\!\top}\!\!\pmx{\widehat M & \widehat N C\\
         B \widehat L & -A+ B\widehat  KC}\pmx{\Id_{n_\xi}\\-\Pi_2}V^{-\top}\!\!\!\!\!\!

@@

evolving in a forward-invariant bounded compact set $\cK\subseteq\cS$ with  $\delta(\eta)=\col(\delta_x(\eta),0,0)$,  $\delta_x(\eta)\in \cC^1$, $|\pderiv{\delta_x}{\eta}(\eta)|<\Delta^\star$  for all $\eta \in \cK$,
converge to an equilibrium $\eta^\star\in \cF^\star_\eta$ with

@@

        \cF^\star_\eta \defeq \{\eta\in\cK: Ax+D\phi(E x) + \delta_x(x)= 0\}.

@@

Nonetheless, Proposition~\ref{prp:2_contr_BMI} presents two main drawbacks. Firstly, it requires imposing matrix inertia constraints, see item iii). A detailed discussion on methods to impose such constraints in the optimization problem is presented in \citep[Section 3]{zoboli2024dynamic}. Secondly, it requires closed-loop trajectories to evolve in a forward invariant bounded compact set $\cK$. This necessity arises from the fact that the proposed result is grounded on the notion of $2$-contraction, which guarantees convergence to nonunique equilibria if the system evolves in forward invariant bounded compact sets \citep[Theorem 2.5]{Michael1995}, \citep[Lemma 6]{cecilia2023generalized}. Therefore, we now present an LMI-based approach allowing the \textit{a-posteriori} verification of such a requirement.

\begin{figure*}
    \begin{equation}
    \label{eqn:ellipsoid_LMI}
    \pmx{\tau_1 P_{\cE_1}- \tau_2 P_{\cE_2} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s - \bX && P_{\cE_1}  -\bY^\top && 0\\\star && - \bZ&&0\\
    \star && \star  && -\tau_1 + \tau_2 }\prec 0.
\end{equation}
    \vspace{.5em}
\hrule
\end{figure*}
\begin{proposition}\label{prp:invariant_set}
    Consider \eqref{eqn:nonlin_sys_closed_loop} and let $\bA = \bA_s +\widetilde \bA$ with $\bA_s$ Hurwitz. Suppose there exist matrices $\bX=\bX^\top$, $\bY$, $\bZ=\bZ^\top$, a set $\cS\subseteq\RR^{n_\eta}$ and a vector $\eta^\star\in\cS$ such that the local sector condition
    \begin{equation}
        \label{eqn:local_sect_eta_reshaped}
        \pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}^\top\pmx{\bX &\bY^\top\\\bY&\bZ}\pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}\le 0
    \end{equation}
    holds for all $\eta\in\cS$, with $\psi(v)\defeq \bD \phi(v)+\widetilde \bA\eta$. 
    If there exist positive scalars $\tau_1$, $\tau_2$, two matrices $P_{\cE_1}\succeq P_{\cE_2}\succ 0$ and a matrix $\Omega\in\RR^{n_\eta \times n_u}$ such that \eqref{eqn:ellipsoid_LMI}\footnotemark[2] holds and such that the ellipsoidal set $\cE_2\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_2} (\eta-\eta^\star) \le 1\} \subseteq\cS$, then $\cE_1\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_1} (\eta-\eta^\star) \le 1\}$ is  attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop} with region of attraction including the  non-empty set $\widetilde\cE\defeq \cE_2\setminus\cE_1$, for all $\eta\in \cS$.
\end{proposition}
The proof of Proposition~\ref{prp:invariant_set} is presented in Section~\ref{sec:proof_prp2}. 
\footnotetext[2]{Given at the top of the next page.}

>>>>

		\author[First]{Andreu Cecilia} 		
        \author[Second]{Samuele Zoboli}
	\author[Second]{Sophie Tarbouriech}

		\address[First]{Universitat Polit\`ecnica de Catalunya, Avinguda Diagonal, 647, 08028 Barcelona, Spain, (e-mail: andreu.cecilia@upc.edu)}
  \address[Second]{LAAS-CNRS, Universit\'e de Toulouse, CNRS, Toulouse, (e-mail: samuele.zoboli@laas.fr, sophie.tarbouriech@laas.fr)}

@@

\section{Problem formulation}
We consider nonlinear systems of the form
\begin{equation}\label{eqn:original_system}
\begin{aligned}
\dot x &=   A x  + D\phi(Ex) + Bu,  \\
 y&= Cx+ G\phi(Ex),
\end{aligned}
\end{equation}
where $x\in\RR^{n_x}$ is the state vector,  $u\in\RR^{n_u}$ the controlled input, $y\in\RR^{n_y}$ the measured outputs and $\phi:\RR^{n_v}\to\RR^{n_v}$ is a sufficiently smooth vector field. Additionally, to recover efficiently solvable conditions, we assume the existence of known matrices $R=R^\top, Q=Q^\top$ and $S$ such that 
\begin{equation}\label{eqn:generic_abstraction}
    \pmx{\Id_{n_x}\\\pderiv{\phi}{x}(v)}^\top\pmx{R&S^\top\\S&Q}\pmx{\Id_{n_x}\\\pderiv{\phi}{x}(v)}\preceq 0 
\end{equation}
holds for all $v\in\cV$ where $\cV\defeq\{v\in\RR^{n_v}:v=E x, x\in\cX\subseteq \RR^{n_x}\}$. We recall that such a class of nonlinearities includes common and interesting family of functions, such as (shifted) monotonic and differentially sector-bounded ones \citep{zoboli2024quadratic}. 
Finally, we assume that in the unforced case, namely when $u=0$, system \eqref{eqn:original_system} presents a set of fixed points $\Omega$. More formally, we assume the existence of a nonempty set of equilibria
\begin{equation}\label{eqn:eq_points}
 \Omega := \{x\in\RR^{n_x} \ | \ Ax+D\phi(E x)= 0\}.
\end{equation}
We highlight that $\Omega$ can include multiple elements, that is, system \eqref{eqn:original_system}  may present multiple equilibrium points for the unforced case $u=0$.

We aim to design a feedback control policy that makes (at least) one point $x^*\in\Omega$ asymptotically stable. In other words, we want to make the set $\Omega$ attractive and forward-invariant.

One could imagine reframing the problem as a standard stabilization one, see e.g. \citep{isidori2017stabilization}. In other words, the problem could be solved by selecting a specific point  $x^*\in\Omega$ and designing a control law making $x^*$ globally asymptotically stable. However, we highlight that the considered problem relaxes the more common stabilization problem in three directions. First, the goal is not to stabilize the system at a specific point in $\Omega$, we only need the system to converge to $\Omega$. Second, the proposed objective allows for the existence of multiple attractive equilibrium points for the closed-loop system.  Third, the set of equilibrium points $\Omega$ may be unknown to the designer. This third point is particularly relevant for feedback policies that include an integral action, e.g. \citep{Astolfi2017,GIACCAGLI2023111106}, which often require an unknown non-zero steady-state control signal to be generated by the integral component.
Another possibility could be to invoice set-stability approaches, e.g. \citep{SHIRIAEV20001709}. However, we remark that $\Omega$ may be disconnected. Moreover, its elements are given by the (possibly unknown) equilibria of the unforced system dynamics. These facts impose additional restrictions that can impact the effectiveness and applicability of set-stability approaches.

Given the abovementioned constraints, in what follows, we present a solution that does not require knowledge of the open-loop equilibria and guarantees their stability via the recent concept of $2$-contraction \citep{Michael1995}. Our approach is based on the recent findings in \citep{ cecilia2023generalized, zoboli2024dynamic} and inspired by the techniques of derivative control and washout filters.

\section{Main result}
\subsection{Proposed approach}
We start by highlighting a necessary property controllers must satisfy to guarantee the open-loop equilibria are preserved by the closed loop.
From \eqref{eqn:original_system}, it is clear that for the closed-loop system to stabilize at  $x^\star\in\Omega$ the control input must vanish at such point.
% dynamic output-feedback law $u=\alpha(y,\xi)$ with state $\xi\in\RR^{n_\xi}$ must satisfy 
% \begin{equation}\label{eqn:noninvasive_cond}
% \alpha(Cx^\star+G\phi(Ex^\star),\xi^\star) = 0, \qquad \forall x^\star \in \Omega,
% \end{equation}
% where $\xi^\star$ is the equilibrium state of the controller corresponding to $x^\star$.
We will denote a control policy that satisfies such a property as \textit{noninvasive}, e.g. \citep{ANGELI2024152}. 
\begin{definition}[Noninvasive controller]
Consider the nonlinear system 
\[
\dot x = f(x,u), \qquad y =h(x),
\]
with $x\in\RR^{n_x}, y\in\RR^{n_y}$ and $u\in\RR^{n_u}$. Let 
\[
\Omega_x\defeq\{x\in\RR^{n_x}:f(x,0)=0\}.
\]
The dynamic output-feedback law $u=\alpha(y,\xi)$  with state $\xi\in\RR^{n_\xi}$ is said to be \textit{ noninvasive} if 
\begin{equation}\label{eqn:noninvasive_cond}
\alpha(h(x^\star),\xi^\star) = 0, \qquad \forall x^\star \in \Omega_x,
\end{equation}
where $\xi^\star$ is the controller equilibrium state associated to $x^\star$.
\end{definition}

In this paper, we propose ensuring such a requirement via integral action. More specifically, we extend the dynamics of \eqref{eqn:original_system} as follows
\begin{equation}\label{eqn:extended_dyn}
\begin{aligned}
\pmx{\dot x\\ \dot z}&= \pmx{A & 0\\0&0} \pmx{x\\ z} + \pmx{B\\\Id_{n_u}} u + \pmx{D\\0} \phi\left(\pmx{E&0} \pmx{x\\ z} \right),  \\
 \pmx{y\\z}&= \pmx{C & 0\\0 & \Id_{n_u}}\pmx{x\\ z} + \pmx{G\\0} \phi\left(\pmx{E&0}\pmx{x\\ z} \right).
\end{aligned}
\end{equation}
At any equilibrium point $(x^\star,z^\star)$ the $z$-dynamics impose $u=0$, and thus $x^\star \in \Omega$.

We remark that such an approach does not allow any of the system's outputs to be regulated. Indeed, from output regulation theory, the number of regulated outputs must be smaller than or equal to the number of inputs, see e.g. \cite{isidori2003robust}. By adding integrator dynamics on all input signals, such a regulation limit is reached.
We also stress that the addition of input-integrator dynamics may cause the loss of controllability properties. This is shown by the following simple example.
Consider the planar linear time invariant system
\begin{equation}
    \label{eqn:ctrlb_example_sys}
    \begin{array}{rl}
     \dot x_1&= x_1+u  \\
     \dot x_2 &= x_1 
\end{array}
\end{equation}
with $x=\col(x_1,x_2)\in\RR^2$ and $u\in\RR$. It can be easily verified that \eqref{eqn:ctrlb_example_sys} is controllable. However, by adding the input-integrator dynamics, the resulting extended system
\[
    \begin{array}{rl}
     \dot x_1&= x_1+u  \\
     \dot x_2 &= x_1 \\
     \dot z &= u,
\end{array}
\]
has a simply stable uncontrollable mode. Luckily, the $2$-contraction framework proposed in this paper only requires a generalized notion of stabilizability properties \citep{cecilia2023generalized}.

@@

holds for all $\eta\in\RR^{n_\eta}$, see \citep[Section 2.5]{zoboli2024dynamic}. 

@@

    D_e \defeq  \pmx{D\\0}, \qquad G_e \defeq  \pmx{G\\0},

@@

        \widehat A &\defeq\pmx{A_eY+B_e \widehat L && A_e + B_e  \widehat K C_e \\ \widehat M && X A_e+ \widehat N  C_e }, \qquad \qquad  
        \widehat D \defeq \pmx{D_e +B_e \widehat K G_e\\ X D_e  + \widehat N G_e}, \qquad \qquad
        \Phi \defeq \pmx{Y & \Id_{n_\xi}\\ \Id_{n_\xi} & X},

@@

   \He{A_e Y+B_e \widehat L} -2\beta_1 Y - \widehat R_{11} && A_e+B_e \widehat K C_e  + \widehat M^\top-2\beta_1 \Id_{n_\xi} -  \alpha YR^\top && D_e +B_e \widehat K G_e -  \alpha YS^\top\\
   \star && \He{ X A_e +  \widehat N  C_e } -2\beta_1 X   -\alpha R&& X D_e  + \widehat N G_e -\alpha S\\
   \star &&\star&& - \alpha Q

@@

    \star && \star& &\He{\Gamma_3 \widehat D} - \alpha  Q

@@

and let \eqref{eqn:generic_abstraction_ctrl} hold for all  $x\in\cX$  with given matrices $R, S, Q$. Select $Z\preceq 0$ such that $Z\preceq R$ and assume there exist $ \widehat R_{11}=\widehat R_{11}^\top,$   scalars $\alpha>0,\beta_0\in\RR$, $\beta_1<0$,  symmetric matrices $X=X^\top, Y=Y^\top \in \RR^{n_\xi\times n_\xi}$, matrices $\widehat K \in \RR^{n_u\times n_y}$, $\widehat L\in \RR^{n_u\times n_\xi}$, $\widehat M\in \RR^{n_\xi\times n_\xi}$, $\widehat N \in \RR^{n_\xi\times n_\xi}$,  matrices $\Gamma_1, \Gamma_2, \Gamma_3$  and $\Sigma=\Sigma^\top\succ 0$  such that 

@@

        \item The matrix $\widehat R \defeq \smallpmat{\widehat R_{11} && \alpha YZ\\ \star && \alpha Z}$ is negative semidefinite;
         \item Inequality \eqref{eqn:2_contraction_BMI}\footnotemark[1] holds with $\widehat S = \smallpmat{  \alpha SY & \alpha S}$ and $\beta_0+\beta_1<0$;

@@

     there exist square invertible matrices $U,V\in \RR^{n_\xi\times n_\xi}$ such that $UV^\top = \Id_{n_\xi}- XY$

@@

         \pmx{N& F}&= U^{-1} (\widehat N - X B \widehat K) ,\\
           L & = (\widehat L - \widehat K C Y) V^{-\top},\\
         M&=U^{-1}\!\pmx{\Id_{n_\xi}\\-X}^{\!\top}\!\!\pmx{\widehat M & \widehat N C\\
         B \widehat L & -A+ B\widehat  KC}\pmx{\Id_{n_\xi}\\-Y}V^{-\top}\!\!\!\!\!\!

@@

evolving in a forward-invariant bounded compact set $\cK\subsetneq \RR^{2 n_\xi}$ with  $\delta(\eta)=\col(\delta_x(\eta),0,0)$,  $\delta_x(\eta)\in \cC^1$, $|\pderiv{\delta_x}{\eta}(\eta)|<\Delta^\star$  for all $\eta \in \cK$,
converge to an equilibrium $\eta^\star\in \Omega_\eta$ with

@@

        \Omega_\eta \defeq \{\eta\in\cK: Ax+D\phi(E x) + \delta_x(x)= 0\}.

@@

Nonetheless, Proposition~\ref{prp:2_contr_BMI} presents two main drawbacks. Firstly, it requires imposing matrix inertia constraints, see item iii). A detailed discussion on methods to impose such constraints in the optimization problem is presented in \citep[Section 3]{zoboli2024dynamic}. Secondly, it requires closed-loop trajectories to evolve in forward invariant bounded compact sets. This necessity arises from the fact that the proposed result is grounded on the notion of $2$-contraction, which guarantees convergence to nonunique equilibria if the system evolves in forward invariant bounded compact sets \citep[Theorem 2.5]{Michael1995}, \citep[Lemma 6]{cecilia2023generalized}. Therefore, we now present an LMI-based approach allowing the \textit{a-posteriori} verification of such a requirement.
% \begin{proposition}\label{prp:invariant_set}
%     Consider \eqref{eqn:nonlin_sys_closed_loop} and assume there exist matrices $\bR=\bR^\top$, $\bS$ and $\bQ$ such that 
%     \begin{equation}\label{eqn:nonincremental_abstraction}
%     \pmx{\eta\\\phi(\bE\eta)}^\top\pmx{\bR&\bS^\top\\\bS&\bQ}\pmx{\eta\\\phi(\bE\eta)}\preceq 0
% \end{equation}
% for all $\eta\in\RR^{n_\eta}$. If there exist scalars $\tau\in\RR$, $\lambda>0$, $\gamma>0$ and a matrix $P_\cE\succ 0$ such that 
% \begin{equation}
%     \label{eqn:ellypsoid_LMI}
%     \pmx{\tau P_\cE+ \bA^\top P_\cE + P_\cE \bA - \gamma \bR & P_\cE \bD -\gamma \bS^\top & 0\\\bD^\top P_\cE -\gamma \bS & -\gamma \bQ&0\\
%     0 & 0 & -\tau \lambda}\preceq 0,
% \end{equation}
% then the ellipsoidal set $\cE=\{\eta\in\RR^{n_\eta}:\eta^\top P_\cE \eta \le \lambda\}$ is forward invariant for \eqref{eqn:nonlin_sys_closed_loop}.
% \end{proposition}

{
\begin{figure*}
    \begin{subequations}\label{eqn:invariance_LMIs}
        \begin{equation}
        \label{eqn:ellypsoid_inclusion}
\lambda_3 \ge \lambda_1, \qquad  P_{\cE_1}- \frac{\lambda_3}{\lambda_2} P_{\cE_2}\succeq 0,
    \end{equation}
\begin{equation}
    \label{eqn:ellypsoid_LMI}
    \pmx{\tau_1 P_{\cE_1}- \tau_2 P_{\cE_2}+ \bA^\top P_{\cE_1} + P_{\cE_1} \bA - \gamma \bR && P_{\cE_1} \bD -\gamma \bS^\top && 0\\\star && -\gamma \bQ&&0\\
    \star && \star  && -\tau_1 \lambda_1+ \tau_2 \lambda_2}\prec 0.
\end{equation}
    \end{subequations}
    \vspace{.5em}
\hrule
\end{figure*}
\begin{proposition}\label{prp:invariant_set}
    Consider \eqref{eqn:nonlin_sys_closed_loop} and assume there exist matrices $\bR=\bR^\top$, $\bS$ and $\bQ$ such that 
    \begin{equation}\label{eqn:nonincremental_abstraction}
    \pmx{\eta\\\phi(\bE\eta)}^\top\pmx{\bR&\bS^\top\\\bS&\bQ}\pmx{\eta\\\phi(\bE\eta)}\preceq 0
\end{equation}
for all $\eta\in\RR^{n_\eta}$. If there exist positive scalars $\tau_1$, $\tau_2$, $\lambda_1$, $\lambda_2$, $\gamma>0$ and two matrices $P_{\cE_1}\succ 0, P_{\cE_2}\succ 0$ such that \eqref{eqn:invariance_LMIs}\footnotemark[2] holds, the ellipsoidal set $\cE_1\defeq\{\eta\in\RR^{n_\eta}:\eta^\top P_{\cE_1} \eta \le \lambda_1\}$ is attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop}, with region of attraction including the set $\cS\defeq\{\eta\in\RR^{n_\eta}: \eta^\top P_{\cE_1}\eta\ge\lambda_1 \land \eta^\top P_{\cE_2}\eta\le\lambda_2\}$.
\end{proposition}
}
The proof of Proposition~\ref{prp:invariant_set} is presented in Section~\ref{sec:proof_prp2}.
\footnotetext[2]{Given at the top of the next page.}

Proposition~\ref{prp:invariant_set} relies on a classical quadratic abstraction \eqref{eqn:nonincremental_abstraction}, while Proposition~\ref{prp:2_contr_BMI} exploits a differential one, i.e. \eqref{eqn:generic_abstraction_ctrl}. In the following technical lemma, we show how classical quadratic abstractions can be recovered from differential ones.

\begin{lemma}
    \label{lem:diff2incr_abstraction}
    If there exist matrices $R_v,S_v, Q_v$ with $Q_v\succ 0$ such that $\phi:\RR^{n_v}\to\RR^{n_v}$ satisfies
    \begin{equation}\label{eqn:diff_abstr}
    \pmx{\Id_{n_v}\\\pderiv{\phi}{v}(v)}^\top\pmx{R_v&S_v^\top\\S_v&Q_v}\pmx{\Id_{n_v}\\\pderiv{\phi}{v}(v)}\preceq 0,
    \end{equation}
    for all $v\in \RR^{n_v}$. Then, 
    \begin{equation}\label{eqn:incr_abstr}
        \pmx{v_1-v_2\\\phi(v_1)-\phi(v_2)}^\top\pmx{R_v&S_v^\top\\S_v&Q_v}\pmx{v_1-v_2\\\phi(v_1)-\phi(v_2)}\preceq 0,
    \end{equation}
    holds for all $v_1,v_2\in\RR^{n_v}$.
\end{lemma}
The proof of Lemma~\ref{lem:diff2incr_abstraction} is presented in Section~\ref{sec:proof_lem1}.

Lemma~\ref{lem:diff2incr_abstraction} shows that, if $\phi(0)=0$, then the selection $v_2 = 0$  and positive definiteness of $\bQ$ allow recovering \eqref{eqn:nonincremental_abstraction} from \eqref{eqn:generic_abstraction_ctrl} by selecting $v_1 = \bE\eta$ and recalling that \eqref{eqn:generic_abstraction_ctrl} can be obtained from inequalities of the form \eqref{eqn:diff_abstr}, see \citep[Section 2.5]{zoboli2024dynamic}.

\begin{remark}
    The results of Proposition~\ref{prp:invariant_set} guarantee the existence of an invariant ellipsoid centered in the origin. However, if the open-loop equilibria are known, such an ellipsoid can be shifted by requiring \eqref{eqn:nonincremental_abstraction} to hold for the error variables $\eta-\eta^\star$ and $\phi(\bE\eta)-\phi(\bE\eta^\star)$. Thanks to the incremental nature of \eqref{eqn:incr_abstr}, Lemma~\ref{lem:diff2incr_abstraction} can still be used to derive the necessary quadratic abstraction from the differential formulation \eqref{eqn:generic_abstraction_ctrl}.
\end{remark}
